% -*-latex-*-
%
%  For more information, please see: http://software.sci.utah.edu
% 
%  The MIT License
% 
%  Copyright (c) 2004 Scientific Computing and Imaging Institute,
%  University of Utah.
% 
%  License for the specific language governing rights and limitations under
%  Permission is hereby granted, free of charge, to any person obtaining a
%  copy of this software and associated documentation files (the "Software"),
%  to deal in the Software without restriction, including without limitation
%  the rights to use, copy, modify, merge, publish, distribute, sublicense,
%  and/or sell copies of the Software, and to permit persons to whom the
%  Software is furnished to do so, subject to the following conditions:
% 
%  The above copyright notice and this permission notice shall be included
%  in all copies or substantial portions of the Software.
% 
%  THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS
%  OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
%  FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL
%  THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
%  LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING
%  FROM, OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER
%  DEALINGS IN THE SOFTWARE.
%

% Document name: package.tex
%

\chapter{Packages}
\label{ch:packages}
\index{packages}

Packages are collections of modules organized by category. \sr{}'s 
packages and their categories are listed below. 

\section{\sr{}}
\label{sec:srpackage}
\index{SCIRun@\sr{}!package}

The \sr{} package provides a base set of datatypes and algorithms.  It
is divided into the following categories:

\begin{itemize}
  \item DataIO
  \item FieldsCreate
  \item FieldsData
  \item FieldsGeometry
  \item FieldsOther
  \item Math
  \item Render
  \item Visualization
\end{itemize}

\section{BioPSE}
\label{sec:biopsepackage}
\index{SCIRun@\BIOPSE{}!package}


The BioPSE package provides modules to solve bioelectric field
problems.  It is divided into the following categories:

\begin{itemize}
\item DataIO
\item Forward
\item Inverse
\item LeadField
\item Modeling
\item Visualization
\end{itemize}

\section{Teem}
\label{sec:teempackage}

Modules in the Teem package encapsulate much of the functionality
found in the \htmladdnormallinkfoot{Teem}
{http://teem.sourceforge.net/} project's, NRRD and Ten libraries.

The NRRD library is used for representing and processing N-dimensional
raster data.  The Ten library is used for diffusion tensor processing,
analysis, and visualization.

It is divided into the following categories:

\begin{itemize}
\item DataIO
\item NrrdData
\item Tend
\item Unu
\end{itemize}

\section{Matlabinterface}
\label{sec:matlabpackage}

The Matlabinterface package provides modules that interact with the
Matlab system.  It is divided into the following categories:

\begin{itemize}
\item DataIO
\end{itemize}

\section{Insight (ITK)}
\label{sec:insightpackage}

The Insight package provides modules that encapsulate
\htmladdnormallinkfoot{ITK toolkit}{http://www.itk.org} filters.

\begin{itemize}
\item Converters
\item DataIO
\item Filters
\end{itemize}

%%\section{DataIO}
%%\label{sec:dataiopackage}


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "usersguide"
%%% End: 
